---
title: "Exoplanet Exploration"
subtitle: "Predicting Future Trends in Exoplanet Exploration"
author: "Annie Li, Matt Clayton, Ethan Wu"
institute: "Bates College"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
library(xaringan)
library(dplyr)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "80%")
```

```{r load-data, include=FALSE}
alldata <- read_csv("/cloud/project/data/filtereddata.csv")
```

```{r discovery-facility-country, include=FALSE}
# Create new variable for country of discovery facility

alldata %>%
  count(disc_facility, sort = TRUE)

alldata <- alldata %>%
  
  mutate(disc_facility = as.character(disc_facility),
         disc_facility_country = case_when(
                                 disc_facility %in% c("Kepler",
                                                       "K2",
                                                       "Transiting Exoplanet Survey Satellite (TESS)",
                                                       "HATNet",
                                                       "HATSouth",
                                                       "W. M. Keck Observatory",
                                                       "XO",
                                                       "KELT-North",
                                                       "KELT",
                                                       "TrES",
                                                       "Acton Sky Portal Observatory",
                                                       "Gemini Observatory",
                                                       "KELT-South",
                                                       "Las Campanas Observatory",
                                                       "MEarth Project",
                                                       "Cerro Tololo Inter-American Observatory",
                                                       "Fred Lawrence Whipple Observatory",
                                                       "Lick Observatory",
                                                       "McDonald Observatory")
                                                        ~ "United States",
                                 disc_facility %in% c("La Silla Observatory",
                                                       "Paranal Observatory",
                                                       "CoRoT")
                                                        ~ "Europe",
                                 disc_facility %in% c("SuperWASP",
                                                       "WASP-South",
                                                       "SuperWASP-North",
                                                       "SuperWASP-South",
                                                       "United Kingdom Infrared Telescope")
                                                       ~ "United Kingdom",
                                 disc_facility %in% c("Roque de los Muchachos Observatory",
                                                       "Calar Alto Observatory")
                                                       ~ "Spain",
                                 disc_facility == "Haute-Provence Observatory" ~ "France",
                                 disc_facility == "Qatar" ~ "Qatar",
                                 disc_facility == "OGLE" ~ "Poland",
                                 disc_facility == "Subaru Telescope" ~ "Japan",
                                 TRUE ~ disc_facility
  ))

```

```{r public-or-private, include=FALSE}

alldata <- alldata %>%
  
  mutate(disc_facility = as.character(disc_facility),
         public_or_private = case_when(
                                 disc_facility %in% c("Kepler",
                                                       "K2",
                                                       "Transiting Exoplanet Survey Satellite (TESS)",
                                                       "HATNet",
                                                       "HATSouth",
                                                       "W. M. Keck Observatory",
                                                       "XO",
                                                       "TrES",
                                                       "Gemini Observatory",
                                                       "Las Campanas Observatory",
                                                       "MEarth Project",
                                                       "Cerro Tololo Inter-American Observatory",
                                                       "Lick Observatory",
                                                       "La Silla Observatory",
                                                       "Paranal Observatory",
                                                       "CoRoT",
                                                       "SuperWASP",
                                                       "WASP-South",
                                                       "SuperWASP-North",
                                                       "SuperWASP-South",
                                                       "United Kingdom Infrared Telescope",
                                                       "Roque de los Muchachos Observatory",
                                                       "Calar Alto Observatory",
                                                       "Haute-Provence Observatory", 
                                                       "Qatar",
                                                       "OGLE",
                                                       "Subaru Telescope")
                                                       ~ "Public",
                                 disc_facility %in% c("KELT-North",
                                                      "KELT",
                                                      "Fred Lawrence Whipple Observatory",
                                                      "McDonald Observatory",
                                                      "Acton Sky Portal Observatory",
                                                      "KELT-South")
                                                      ~ "Private",
                                 TRUE ~ disc_facility
  ))

```

```{r telescope-type, include=FALSE}

alldata <- alldata %>%
  
  mutate(disc_facility = as.character(disc_facility),
         facility_type = case_when(
                                 disc_facility %in% c( "HATNet",
                                                       "HATSouth",
                                                       "W. M. Keck Observatory",
                                                       "XO",
                                                       "TrES",
                                                       "Gemini Observatory",
                                                       "Las Campanas Observatory",
                                                       "MEarth Project",
                                                       "Cerro Tololo Inter-American Observatory",
                                                       "Lick Observatory",
                                                       "La Silla Observatory",
                                                       "Paranal Observatory",
                                                       "SuperWASP",
                                                       "WASP-South",
                                                       "SuperWASP-South",
                                                       "SuperWASP-North",
                                                       "United Kingdom Infrared Telescope",
                                                       "Roque de los Muchachos Observatory",
                                                       "Calar Alto Observatory",
                                                       "Haute-Provence Observatory", 
                                                       "Qatar",
                                                       "OGLE",
                                                       "Subaru Telescope",
                                                       "KELT-North",
                                                       "KELT",
                                                       "Fred Lawrence Whipple Observatory",
                                                       "McDonald Observatory",
                                                       "Acton Sky Portal Observatory",
                                                       "KELT-South")
                                                       ~ "Observatory",
                                 disc_facility %in% c("CoRoT",
                                                      "Kepler",
                                                      "K2",
                                                      "Transiting Exoplanet Survey Satellite (TESS)")
                                                      ~ "Space Telescope",
                                 TRUE ~ disc_facility
  ))
```

```{r facility-latitude, include=FALSE}

alldata <- alldata %>%
  
  mutate(disc_facility = as.character(disc_facility),
         facility_latitude = case_when(
                                 disc_facility == "HATNet" ~ 31.674920,
                                 disc_facility == "HATSouth" ~ -29.014600,
                                 disc_facility == "W. M. Keck Observatory" ~ 19.826360,
                                 disc_facility == "XO" ~ 20.708300,
                                 disc_facility == "TrES" ~ 35.1210,
                                 disc_facility == "Gemini Observatory" ~ 19.823960,
                                 disc_facility == "Las Campanas Observatory" ~ -29.014600,
                                 disc_facility == "MEarth Project" ~ 31.674920,
                                 disc_facility == "Cerro Tololo Inter-American Observatory" ~ -30.169660,
                                 disc_facility == "Lick Observatory" ~ 37.341373,
                                 disc_facility == "La Silla Observatory" ~ -29.261170,
                                 disc_facility == "Paranal Observatory" ~ -24.627220,
                                 disc_facility == "SuperWASP" ~ 28.4537,
                                 disc_facility == "WASP-South" ~ -26.030260,
                                 disc_facility == "SuperWASP-North" ~ 28.75651,
                                 disc_facility == "SuperWASP-South" ~ -26.030260,
                                 disc_facility == "United Kingdom Infrared Telescope" ~ 19.822500,
                                 disc_facility == "Roque de los Muchachos Observatory" ~ 28.4522,
                                 disc_facility == "Calar Alto Observatory" ~ 37.220790,
                                 disc_facility == "Haute-Provence Observatory" ~ 43.930830, 
                                 disc_facility == "Qatar" ~ 32.902359,
                                 disc_facility == "OGLE" ~ -29.0036,
                                 disc_facility == "Subaru Telescope" ~ 19.701610,
                                 disc_facility == "KELT-North" ~ 31.71272,
                                 disc_facility == "KELT" ~ 31.71272,
                                 disc_facility == "Fred Lawrence Whipple Observatory" ~ 31.674920,
                                 disc_facility == "McDonald Observatory" ~ 30.679770,
                                 disc_facility == "Acton Sky Portal Observatory" ~ 42.477587,
                                 disc_facility == "KELT-South" ~ 31.71272,
                                 disc_facility == "CoRoT" ~ 45.592759,
                                 disc_facility == "Kepler" ~ 28.446671,
                                 disc_facility == "K2" ~ 28.446671,
                                 disc_facility == "Transiting Exoplanet Survey Satellite (TESS)" ~ 28.562110,
                                                     
                                 TRUE ~ 0
  ))
```

```{r facility-longitude, include=FALSE}

alldata <- alldata %>%
  
  mutate(disc_facility = as.character(disc_facility),
         facility_longitude = case_when(
                                 disc_facility == "HATNet" ~ -110.949600,
                                 disc_facility == "HATSouth" ~ -70.692600,
                                 disc_facility == "W. M. Keck Observatory" ~ -155.475000,
                                 disc_facility == "XO" ~ -156.257100,
                                 disc_facility == "TrES" ~ -111.3952,
                                 disc_facility == "Gemini Observatory" ~ -155.469800,
                                 disc_facility == "Las Campanas Observatory" ~ -70.692600,
                                 disc_facility == "MEarth Project" ~ -110.949600,
                                 disc_facility == "Cerro Tololo Inter-American Observatory" ~ -70.806530,
                                 disc_facility == "Lick Observatory" ~ -121.642944,
                                 disc_facility == "La Silla Observatory" ~ -70.731330,
                                 disc_facility == "Paranal Observatory" ~ -70.404170,
                                 disc_facility == "SuperWASP" ~ -17.5245,
                                 disc_facility == "WASP-South" ~ 27.924900,
                                 disc_facility == "SuperWASP-North" ~ -17.89163,
                                 disc_facility == "SuperWASP-South" ~ 27.924900,
                                 disc_facility == "United Kingdom Infrared Telescope" ~ -155.470700,
                                 disc_facility == "Roque de los Muchachos Observatory" ~ -17.5330,
                                 disc_facility == "Calar Alto Observatory" ~ -2.546575,
                                 disc_facility == "Haute-Provence Observatory" ~ 5.713333, 
                                 disc_facility == "Qatar" ~ -105.530205,
                                 disc_facility == "OGLE" ~ -70.4156,
                                 disc_facility == "Subaru Telescope" ~ -155.089600,
                                 disc_facility == "KELT-North" ~ -110.60182,
                                 disc_facility == "KELT" ~ -110.60182,
                                 disc_facility == "Fred Lawrence Whipple Observatory" ~ -110.949600,
                                 disc_facility == "McDonald Observatory" ~ -104.024800,
                                 disc_facility == "Acton Sky Portal Observatory" ~ -71.453447,
                                 disc_facility == "KELT-South" ~ -110.60182,
                                 disc_facility == "CoRoT" ~ 63.333059,
                                 disc_facility == "Kepler" ~ -80.566109,
                                 disc_facility == "K2" ~ -80.566109,
                                 disc_facility == "Transiting Exoplanet Survey Satellite (TESS)" ~ -80.577180,
                                                     
                                 TRUE ~ 0
  ))
```


```{r}
style_xaringan(
  title_slide_background_image = "img/SPACE_2.jpg"
)
```

### Plotting facilities and discoveries
```{r construct-facility_df}
facility_df <- alldata %>% 
  group_by(disc_facility) %>%
  mutate(count = n()) %>%
  select(disc_facility, facility_type, facility_latitude, facility_longitude, count) %>%
  distinct()
```


class: center, middle

## A statement of the overall goal / research question

---

class: inverse, center, middle

# Section title

---

# Hello World

- Click the `Knit` button to compile your presentation

- Make sure to commit and push all resulting files to your GitHub repo

---

class: inverse, middle, center

# Using xaringan

---

# xaringan

- The presentation is created using the `xaringan` package

- Use `---` to separate slides and `--` for incremental builds

--

- Like this

---

# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
# ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
#  geom_point() + 
#  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

```{r discovery-method-over-time, echo = FALSE}
alldata %>%
  ggplot(mapping = aes(x = disc_year, fill = discoverymethod)) +
  geom_bar() +
  labs(
        x = "Distance of Discovery (Parsecs)",
        y = "Number of exoplanets", 
        title = "Number of exoplanets discovered per year",
        subtitle="By discovery method") 
```

---
# Plots
```{r discovery-method-vs-planet_radius-and-planet-mass, echo = FALSE, fig.show="hold", out.width="50%"}
alldata %>%
  ggplot(mapping = aes(x = pl_radius,
                       y = sy_distance,
                       color = discoverymethod)) +
  geom_jitter() +
  coord_flip() +
  labs(
        x = "Distance of Discovery (Parsecs)",
        y = "Planet radius [Earth radius]", 
        title = "Planet radius and distance of discovery",
        subtitle="By discovery method" ) +
  theme(legend.position = "bottom") +
  theme(legend.title = element_blank())

alldata %>%
  ggplot(mapping = aes(x = pl_mass,
                       y = sy_distance,
                       color = discoverymethod)) +
  geom_jitter() + 
  coord_flip() +
  labs(
        x = "Distance of Discovery (Parsecs)",
        y = "Planet mass [Earth mass]", 
        title = "Planet mass and distance of discovery",
        subtitle="By discovery method" ) +
  theme(legend.position = "bottom") +
  theme(legend.title = element_blank())
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
